\input{../header.tex}

\begin{document}
\title{Problem Set 2}
\date{}
\maketitle

\paragraph*{1. Generalized Bernstein-Vazirani}
Let $M,n$ be a positive integers.  Let $s \in \Z_M^n$ and define the 
function $f_s : \Z_M^n \rightarrow \Z_M$ by $f_s(x) = \langle s, x \rangle \mod M$.  
Given access to an oracle $O_{f_s}$ which for $x \in \Z_M^n , b \in \Z_M$ acts as 
$O_{f_s} \ket{x} \ket{b} = \ket{x} \ket{b + f_s(x) \bmod M}$, design a 
quantum algorithm that computes $s$ with one application of $O_{f_s}$.  

Hint: You may want to generalize the ``phase-kickback trick'' to show with the oracle 
$O_{f_s}$ you can also implement an oracle $O'_{f_s}$ with the behavior
\[
O'_{f_s} \ket{x}\ket{b} = \omega^{-f_s(x) \cdot b} \ket{x}\ket{b} 
\]
where $\omega = e^{2\pi \iu/M}$.  

Bonus: What kind of errors in the oracle can your algorithm tolerate (analogous to what we saw in problem 7 of 
problem set 1)?

\paragraph*{2. Continued fractions}
In the classical post-processing of Shor's period finding algorithm we have a fraction $b/N$ and want to find the 
best rational approximation to this number whose denominator is at most $M$.  In lecture we said this can be 
done in polynomial time as the task can be written as a two-variable integer linear program.  Now we see a direct 
way to do this via continued fraction expansion.  A nice discussion of continued fractions, including all the material 
below, can be found in Chapter 10 of Hardy and Wright's An introduction to the theory of numbers.

A finite continued fraction is an expression of the form
\[
a_0 + \frac{1}{a_1 + \frac{1}{a_2 + \frac{1}{a_3 + \frac{1}{\cdots + \frac{1}{a_t}}}}} \enspace.
\]
We will denote this number by $[a_0, \ldots, a_t]$.  For $0 \le j \le t$ we call $[a_0, \ldots, a_j]$ the 
$\jth$ convergent to $[a_0, \ldots, a_p]$.  A continued fraction $[a_0, \ldots, a_t]$ is called \emph{simple}
if $a_1, \ldots, a_p$ are all positive integers ($a_0$ can be non-positive).  Every rational number can be represented by a finite simple 
continued fraction.

Here is an algorithm to find such a representation.  Let $x$ be a positive rational number.  Then set 
\begin{align*}
a_0 = \floor{x}, &\quad x_1 = \frac{1}{x-a_0} \\
a_1 = \floor{x_1}, &\quad x_2 = \frac{1}{x_1-a_1} \\
a_2 = \floor{x_2}, &\quad x_3 = \frac{1}{x_2-a_2} \\
&\cdots
\end{align*}
The essential principle at work here is that $x = a_0 + \frac{1}{a_1'}$ where $a_1' = \frac{1}{x-a_0}$.  Then 
since $[a_0, [a_1, \ldots, a_t]] = [a_0, a_1, \ldots, a_t]$ our task becomes to find a continued fraction expansion 
of $a_1'$ which we do by the same procedure.  

One can also find an inductive expression for $[a_0, \ldots, a_j]$.  If 
\begin{align*}
p_0 &= a_0, & p_1 &= a_1 a_0 + 1, & p_j &= a_j p_{j-1} + p_{j-2} \\
q_0 &= 1,  & q_1 &= a_1, & q_j &= a_j q_{j-1} + q_{j-2}
\end{align*}  
then $[a_0, \ldots, a_j] = \frac{p_j}{q_j}$ and this is in lowest terms.  Note that $q_j \ge 2 q_{j-2}$ thus $q_j$ increases 
at least exponentially.  An important property of the continued fraction expansion for the application in Shor's algorithm 
is that if 
\[
|x - \frac{c}{d}| \le |x - \frac{p_j}{q_j}|
\]
then $d \ge q_j$.  

Now the questions:
\begin{enumerate}
\item Find the continued fraction expansion of $\frac{527}{1024}$.  
\item Look at the $\jth$ convergents of your expression and 
make a conjecture about the even and odd numbered convergents (you do not need to prove it).
\item (Optional but could be helpful for Problem 3) Write a program in any language to compute 
a continued fraction of an input number up to a given accuracy.
\end{enumerate}

\paragraph*{3. Factoring 21}
Let's factor the number $M=21$ using Shor's algorithm.
\begin{enumerate}
  \item List all numbers in $\Z_{21}$ that are relatively prime to $21$.  These are the elements of the multiplicative group $\Z_{21}^\times$.  Compute 
  the order $\ord_{21}(x)$ of all elements in $\Z_{21}^\times$.
  \item Recall that in Shor's algorithm we want to find an $x$ of even order $d$ such that $x^{d/2} \ne -1 \mod M$.  Call such an $x$ \emph{good}. 
  Identify all the good $x \in \Z_{21}^\times$ with $\ord_{21}(x) = 6$ and for these verify that $\gcd(x^3 \pm 1, 21)$ gives a nontrivial factor of $21$.
  \item Choose a good $x$ of order $6$ from the previous step.  Now let's simulate finding the period of $f(j) = x^j \bmod{21}$.  Using the Octave FTperiod program 
  \footnote{Currently I have only added the sampling functionality to the Matlab/Octave program.  If I have time I will also add it to the python version.  
  Octave programs can be run online at \url{https://octave-online.net/}.}
  \url{https://github.com/troyjlee/qalgo/tree/main/CODE} with $N = 21^2, s=6$.  This simulates randomly sampling a state $\ket{g_t}$ and 
  measuring $F_N \ket{g_t}$ to see an index $b$.  Use continued fraction expansion on $b/N$ and see if you can recover $\ord_{21}(x)$.  
  It may take several attempts.  Record the values you see and how many attempts it takes.
\end{enumerate}

\paragraph*{4. Assumptions} 
Where in the proof of correctness of Shor's algorithm for the general period finding problem with a function 
$f: \Z_N \rightarrow [M]$ do we use the assumption that $N > M^2/2$?  What can go wrong without this assumption?

\paragraph*{5. Cosets}
Let $G$ be a finite group and $K, L \le G$ subgroups of $G$.  For $a,b \in G$ let $aK = \{a \cdot k : k \in K\}$ 
be a left coset of $K$ and $bL$ similarly be a left coset of $L$.  If $d = |K \cap L|$ show that $|aK \cap bL| \in \{0,d\}$.  

\paragraph*{6. Finding all ones}
Let $N = 2^n$ and $x \in \{0,1\}^N$ and \emph{assume you know} that $x$ has $k$ many ones.  
\begin{enumerate}
\item In lecture we showed how to find an $i \in N$ such that $x_i =1$ with constant probability by a quantum algorithm after $O(\sqrt{N/k})$ many queries to $x$.  
Show how to boost this success probability to $1-1/N^2$ using $O(\sqrt{N/k} \log(N))$ many queries to $x$.
\item Give a quantum algorithm  to find \emph{all} the ones in $x$ with constant probability after $O(\sqrt{kN} \log(N))$ many queries to $x$.
\end{enumerate}

\paragraph*{7. Exact searching}
Do Exercise~4 in Chapter~7 of Ronald de Wolf's lecture notes \url{https://arxiv.org/abs/1907.09415}.  For part~(c) you may assume 
you have access to the phase oracle $O_{f,\pm}$ for $f$ and may use extra ancillas and any elementary gates you like.
\end{document}
















